---
title: "User Guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{User Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r setup, include = FALSE}
key_available <- httr2::secret_has_key("DTMAPIR_KEY")

if (key_available) {
  Sys.setenv(
    BASELINE_DTM_SUBSCRIPTION_KEY = httr2::secret_decrypt(
      "a6Duw21vazOwY2uvRjZKcNhsCrqkFKNKZLjpveBU55Ya6pmbPIwjRfYD6QM30aoy",
      "DTMAPIR_KEY"
    ),
    HNA_DTM_SUBSCRIPTION_KEY = httr2::secret_decrypt(
      "7ilgaf5ZSF9sqM2A-3RQgji0kiAfkqgCx1JIwbrMZKPayEtc6979ZCAx7qPD7ghD",
      "DTMAPIR_KEY"
    )
  )
}

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  error = TRUE,
  eval = key_available
)
```

```{r eval = !key_available, echo = FALSE, comment = NA}
message("No token available. Code chunks will not be evaluated.")
```

## Introduction

The `dtmapi` package provides functions to interact with the Displacement Tracking Matrix (DTM) API. This vignette demonstrates how to use the package's functions to fetch data from the API.

Please note that despite the overarching term DTM API, there are,
in fact, two distinct APIs which are currently available. These are the Baseline API (which is what simply "the DTM API" *used* to refer to) and the Humanitarian Needs Assessment (HNA) API.

Reflecting this, this R package comes with a set of functions that are clearly named so as to distinguish between these two APIs. Those prefixed with `baseline_` and those prefixed with `hna_`.

The functions covered include:

-   `baseline_get_countries()`

-   `baseline_get_operations()`

-   `baseline_get_admin0()`

-   `baseline_get_admin1()`

-   `baseline_get_admin2()`

-   `hna_get_countries()`

-   `hna_get_catalog()`

-   `hna_get_admin2()`

As well as these functions which are concern your API subscription keys.

- `baseline_set_subscription_key()`

- `baseline_get_subscription_key()`

- `hna_set_subscription_key()`

- `hna_get_subscription_key()`

Notice that as a consequence of there being two distinct APIs, each API has its own pair of subscription key-related functions. Their usage is covered in this guide.

## Install Package

The latest version of the `dtmapi` package can be installed directly from [GitHub](https://github.com/Displacement-Tracking-Matrix/dtmapi-R) using the following command (make sure the `remotes` package is installed):
```{r install, eval = FALSE}
remotes::install_github("Displacement-Tracking-Matrix/dtmapi-R")
```

## Load Package

After installation, load the package using library():
```{r import, eval = TRUE}
library(dtmapi)
```

## Setting the Subscription Keys
After creating a subscription key on https://dtm-apim-portal.iom.int, that key
needs to be given to R. The package has two sets of functions, each with its own subscription key: the `baseline_` functions use the `BASELINE_DTM_SUBSCRIPTION_KEY` environment variable, and the `hna_` functions use the `HNA_DTM_SUBSCRIPTION_KEY` environment variable. There are three main ways to set these, as described below. The examples use the Baseline key; the HNA key works the same way, with `hna_set_subscription_key()` and `HNA_DTM_SUBSCRIPTION_KEY` in place of `baseline_set_subscription_key()` and `BASELINE_DTM_SUBSCRIPTION_KEY`.

### 1. Using `baseline_set_subscription_key()`: interactive, safe option

By calling `baseline_set_subscription_key()`, after which the user is prompted to input the key into a pop-up field.
```{r set_subscription_key_demo, eval = FALSE}
baseline_set_subscription_key()
```

Note that if the environment is not interactive, it is likely for the user to receive an error message instead, such as in the case of using Google Colab. In such cases see the other options described further below.

### 2. Using `baseline_set_subscription_key()`: the non-interactive, discouraged option

If the interactive option above fails for the user, such that there is an error instead of a pop-up, the user may rely on passing their key directly into `baseline_set_subscription_key()` by naming the `key` parameter.
```{r, eval = FALSE}
# Specifying `key` is required for non-interactive use.
baseline_set_subscription_key(key = "mysubscriptionkey")
```

This is basically using `baseline_set_subscription_key()` as a wrapper for setting up the environment variable `BASELINE_DTM_SUBSCRIPTION_KEY` using base R's `Sys.setenv()`, as shown below:
```{r, eval = FALSE}
Sys.setenv("BASELINE_DTM_SUBSCRIPTION_KEY" = "mysubscriptionkey")
```

Please note that by doing this in, say, a Google Colab script, .R script, or Jupyter notebook, you are leaving a secret (your subscription key) exposed and available for anyone to read with ease. This is why this option is not recommended, but it will allow you to get things working faster if you are not familiar with managing secrets in a non-interactive context.

### 3. Setting up the environment variable with good secrets management

If one would like to avoid re-inputting the subscription key for each R session, and they would like to set the subscription key in a way that is both non-interactive and safer, then this becomes a matter of knowing the right secrets management practices depending on where R is being used (e.g. whether that is in a local desktop, a cloud-based programming environment, a server, etc.).

In the vast majority of cases, the approach is the fundamentally the same: anything that directly stores your key should be **separate from the rest of your code**, and your code accesses the key through an **environment variable**. That is the main principle to the most common way to manage your secrets. What differs by where R is being used is _how_ this principle is applied. Below, the user guide will mostly focus on how this is handled when using a code editor such as VSCode, RStudio, or Positron, as this is the most common case. If these do not reflect your circumstance or use-case, then researching how to manage secrets generally for that circumstance should suffice, as the only thing specific about doing so for the `dtmapi` package is just the names of the environment variables, `BASELINE_DTM_SUBSCRIPTION_KEY` and `HNA_DTM_SUBSCRIPTION_KEY`.

**Using a code editor such as VSCode or RStudio on a desktop:**

Use a `.Renviron` file to set the keys. You can create it from scratch, and its contents should look something like this:
```
BASELINE_DTM_SUBSCRIPTION_KEY = "mysubscriptionkey"
HNA_DTM_SUBSCRIPTION_KEY = "myothersubscriptionkey"
Some_Other_Env_Var = "something else if desired"
```
For this to work robustly, your `.Renviron` should be in the same *working directory* that you would like to use for your R sessions (i.e. it should be in the same location or folder that you would like your R sessions to use as a default). You can see which folder that currently is by running:

```{r, eval = FALSE}
getwd()
```

This is important to know because, if your R session can start in the same working directory that the `.Renviron` file is located in, then the environment variables specified there are automatically loaded for your R session!

Usually, your choice of working directory is best managed by creating an _R project_ or equivalent (such as _VS Code workspaces_). That is to say, one should have a dedicated folder for their specific project, the R sessions they run should make that folder the working directory, and `.Renviron` should be placed there as well. For more information, depending on whether you use RStudio, VSCode, or Positron, see: [Using RStudio Projects](https://support.posit.co/hc/en-us/articles/200526207-Using-RStudio-Projects), [What is a VS Code Workspace?](https://code.visualstudio.com/docs/editing/workspaces), or [Launching Positron in a workspace](https://positron.posit.co/migrate-rstudio-rproj.html#launching-positron-in-a-workspace).

If one is unfamiliar and struggling to create the .Renviron file, then one quick way to proceed with editing the `.Renviron` file or placing it in their project directory (if they have one) is by calling the following line of code (make sure the `usethis` package is installed):
```{r r_environ_demo_project, eval = FALSE}
usethis::edit_r_environ("project")
```

Or, if one is not using an R project or equivalent, then this is an option as well, though using the general user directory is not recommended as it is likely to get lost:
```{r r_environ_demo_user, eval = FALSE}
usethis::edit_r_environ("user")
```

Remember: NEVER SHARE THE `.RENVIRON` FILE.

## Get All Countries

The `baseline_get_countries()` function retrieves a list of all countries from the DTM API.

```{r get_country}
# Fetch all countries
countries_df <- baseline_get_countries()

# Display the first few rows of the data frame
head(countries_df)
```

## Get All Operations

The `baseline_get_operations()` function retrieves a list of all operations from the DTM API.

```{r get_operations}
# Fetch all operations
operations_df <- baseline_get_operations()

# Display the first few rows of the data frame
head(operations_df)
```

## Get IDP Data at Admin Level 0

The `baseline_get_admin0()` function retrieves Internally Displaced Persons (IDP) data aggregated at the country level.

```{r get_idp_admin0}
# Fetch IDP data at Admin Level 0
idp_admin0_df <-
  baseline_get_admin0(country_name = "Ethiopia",
                      from_round_number = 0,
                      to_round_number = 10)

# Display the first few rows of the data frame
head(idp_admin0_df)
```

## Get IDP Data at Admin Level 1

The `baseline_get_admin1()` function retrieves IDP data aggregated at Admin Level 1.

```{r get_idp_admin1}
# Fetch IDP data at Admin Level 1
idp_admin1_df <-
  baseline_get_admin1(country_name = "Sudan",
                      admin1name = "Blue Nile",
                      from_reporting_date = "2020-01-01",
                      to_reporting_date = "2024-08-15")

# Display the first few rows of the data frame
head(idp_admin1_df)
```

## Get IDP Data at Admin Level 2

The `baseline_get_admin2()` function retrieves IDP data aggregated at Admin Level 2.
```{r get_idp_admin2}
# Fetch IDP data at Admin Level 2
idp_admin2_df <-
  baseline_get_admin2(operation = "Displacement due to conflict",
                      country_name = "Lebanon")

# Display the first few rows of the data frame
head(idp_admin2_df)
```

## Get All Countries (HNA)

The `hna_get_countries()` function retrieves a list of all countries covered by the DTM Household Needs Assessment (HNA) API.

```{r hna_get_country}
# Fetch all HNA countries
hna_countries_df <- hna_get_countries()

# Display the first few rows of the data frame
head(hna_countries_df)
```

## Get the HNA Data Catalog

The `hna_get_catalog()` function retrieves the HNA data catalog, which provides information on the indicators available in the HNA API, including their name, description, and data type.

```{r hna_get_catalog}
# Fetch the HNA data catalog
hna_catalog_df <- hna_get_catalog()

# Display the first few rows of the data frame
head(hna_catalog_df)
```

## Get HNA Data at Admin Level 2

The `hna_get_admin2()` function retrieves HNA data at Admin Level 2. Unlike the other functions, it returns a list of two elements: `data`, a data frame holding the requested data, and `pagination`, a list containing metadata on pagination.

```{r hna_get_admin2}
# Fetch HNA data at Admin Level 2
hna_admin2 <-
  hna_get_admin2(admin0name = "Nigeria",
                 year = 2025)

# Display the first few rows of the data
head(hna_admin2$data)

# Display the pagination metadata
hna_admin2$pagination
```

## Function Arguments

Here are the descriptions for the arguments used in the functions of the `dtmapi` package to get data.

### Baseline Arguments

For the `baseline_` functions that get IDP data, at least one of the following parameters must be provided: operation, country_name, or admin0pcode.

- **operation**: Optional; Name of the DTM operation for which the data was collected.

- **country_name**: Optional; Name of the country where the data was collected.

- **admin0pcode**: Optional; Country code (ISO 3166-1 alpha-3).

- **admin1name**: Optional; Name of level 1 administrative boundaries.

- **admin1pcode**: Optional; Place code of level 1 administrative boundaries.

- **admin2name**: Optional; Name of level 2 administrative boundaries.

- **admin2pcode**: Optional; Place code of level 2 administrative boundaries.

- **from_reporting_date**: Optional; Start date for the reporting period (format: 'YYYY-MM-DD').

- **to_reporting_date**: Optional; End date for the reporting period (format: 'YYYY-MM-DD').

- **from_round_number**: Optional; Starting round number for the data collection range.

- **to_round_number**: Optional; Ending round number for the data collection range.

### HNA Arguments

For `hna_get_admin2()`, at least one of the following parameters must be provided: admin0name or admin0pcode.

- **admin0name**: Optional; Name of the country where the data was collected.

- **admin0pcode**: Optional; Country code (ISO 3166-1 alpha-3).

- **population_group**: Optional; A specific subpopulation (e.g. internally displaced persons).

- **year**: Optional; Year of data collection.

- **page**: Optional; Pagination, to fetch a certain chunk of the data.
